Accessible and Completed Datasets yet Majority are Bronze-Graded and not Updated Regularly with Missing Metadata

### 2025-05-15 paper/paper.qmd
Updated intro
Typo fixed

In this paper, we analyzed data provided by Open Data Toronto on the data quality grade of the datasets in their catalogue and their characteristics like "accessibility", "completeness", "freshness", "metadata", and "usability". Farrow analyzed the data quality scores of the different datasets on the Open Data Toronto portal and found that the metadata and freshness scores were poor overall [@citeFarrow]. However there has been no analysis done on the scores and how they compare for different graded datasets as of 2025. In our findings, all datasets in Open Data Toronto portal are accessible and a majority of them having minimial missing data and are usable. It also showed that 56% of their datasets are graded bronze with bronze-graded datasets being less likely to be updated and have completed metadata fields. These bronze-graded datasets contribute the poor metadata and freshness scores we saw. These findings can inform Open Data Toronto about specific graded datasets that should be given more attention and the specific qualities of them such as metadata completion, which impacts the quality of the data used in the media and in policymaking. These findings can also be informative for users of the portal in order to understand what goes behind the grade of datasets on the portal and what they mean.

@@

For more technical details about how each dimension is calculated for a dataset and the final data quality score, it can be through the following link:[https://github.com/open-data-toronto/framework-data-quality/blob/master/data_quality_score.ipynb](https://github.com/open-data-toronto/framework-data-quality/blob/master/data_quality_score.ipynb) [@citeDQSNotebook]. However, it is worth noting that each dimension are calculated based on data obtained on the metadata information and the dataset itself. If the data quality score of a dataset has a normalized score of less than 60%, they are given the grade "Bronze", if the normalized score is between 60% and 80%, it is given the grade "Silver", and finally if the normalized score is over 80%, tit is given the grade "Gold". The Information & Technology department at Open Data Toronto collected datasets on their portal using the CKAN Datastore API in order to give each of them a data quality score and grade [@citeCatalogue; @citeCatalogueSteps]. The score for each dataset in the portal is recalculated every week by the team [@citeCatalogue].

>>>>

In this paper, we analyzed data provided by Open Data Toronto on the data quality grade of the datasets in their catalogue and their characteristics like "accessibility", "completeness", "freshness", "metadata", and "usability". Farrow analyzed the data quality scores of the different datasets on the Open Data Toronto portal and found that the metadata and freshness scores were poor overall [@citeFarrow]. However there has been no analysis done on the characteristics and how they compare for different graded datasets as of 2025. In our findings, all datasets in Open Data Toronto portal are accessible and a majority of them having minimial missing data and are usable. It also showed that 56% of their datasets are graded bronze with bronze-graded datasets being less likely to be updated and have completed metadata fields. These bronze-graded datasets contribute the poor metadata and freshness scores we saw. These findings can inform Open Data Toronto about specific graded datasets that should be given more attention and the specific qualities of them such as metadata completion, which impacts the quality of the data used in the media and in policymaking. These findings can also be informative for users of the portal in order to understand what goes behind the grade of datasets on the portal and what they mean.

@@

For more technical details about how each dimension is calculated for a dataset and the final data quality score, it can be through the following link:[https://github.com/open-data-toronto/framework-data-quality/blob/master/data_quality_score.ipynb](https://github.com/open-data-toronto/framework-data-quality/blob/master/data_quality_score.ipynb) [@citeDQSNotebook]. However, it is worth noting that each dimension are calculated based on data obtained on the metadata information and the dataset itself. If the data quality score of a dataset has a normalized score of less than 60%, they are given the grade "Bronze", if the normalized score is between 60% and 80%, it is given the grade "Silver", and finally if the normalized score is over 80%, it is given the grade "Gold". The Information & Technology department at Open Data Toronto collected datasets on their portal using the CKAN Datastore API in order to give each of them a data quality score and grade [@citeCatalogue; @citeCatalogueSteps]. The score for each dataset in the portal is recalculated every week by the team [@citeCatalogue].
